---
title: "SentimentAnalysisProject"
author: "Condag, Gagante, Gonzaga"
date: "10/12/2024"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(dplyr)
library(syuzhet)
library(ggplot2)
library(tidyverse)
library(stringr)
library(readr)
library(lubridate)

iniTweets <- read.csv("tweetsDF.csv")

tweetsDf <- iniTweets

```
---


--- Cleaning Data ---

```{r}
cleanTweets <- tweetsDf %>%
  mutate(text = str_to_lower(text),                     
         text = str_remove_all(text, "http\\S+"),
         text = str_remove_all(text, "[^\\w\\s]"),
         text = str_remove_all(text, "#\\w+"),
         text = str_remove_all(text, "@\\w+"),          
         text = str_remove_all(text, "\\d+"),    
         text = str_squish(text),                      
         sentiment = get_sentiment(text, method = "bing"))

write.csv(cleanTweets, "cleaned_tweets.csv", row.names = FALSE)

```
---



--- Trend Analysis ---
```{r}
cleanTweets <- cleanTweets %>%
  mutate(date = ymd_hms(created)) %>%  
  mutate(hour = hour(date))

byHour <- cleanTweets %>%
  group_by(day = as.Date(date), hour) %>%
  summarise(tweet_count = n(), .groups = "drop")

day1 <- byHour %>% filter(day == unique(day)[1])
day2 <- byHour %>% filter(day == unique(day)[2])
day3 <- byHour %>% filter(day == unique(day)[3])

all_days <- bind_rows(
  day1 %>% mutate(day_label = paste("Day", 1)),
  day2 %>% mutate(day_label = paste("Day", 2)),
  day3 %>% mutate(day_label = paste("Day", 3))
)


ggplot(all_days, aes(x = hour, y = tweet_count, group = day, color = as.factor(day))) +
  geom_line(size = 1) +
  geom_point() +
  scale_x_continuous(breaks = seq(0, 23, by = 1)) + 
  labs(
    title = "Tweet Counts by Hour",
    x = "Hour of the Day",
    y = "Number of Tweets",
    color = "Date"
  ) +
  theme_minimal()

```

The graph shows Hourly Tweet Counts from Day 1 to Day 3, with tweet data represented across different times of the day. 

- In Day 1, The red line remains consistently flat at the bottom, indicating almost no tweeting activity throughout the 24 hours.

- In Day 2, The green line shows no activity until around 14:00 (2 PM), after which there is a sharp increase in tweet counts. The number of tweets spikes significantly between 14:00 and 18:00.  Afterward, the tweet count drops but remains moderately high until the end of the day.

- In Day 3, The blue line shows consistent tweeting activity throughout the day, starting with a high volume (around 2000-3000 tweets) during the early morning hours (0:00 - 5:00). The tweet count gradually decreases throughout the morning and early afternoon. There is a slight dip and, followed by a minor uptick before declining again in the late afternoon.

#INSIGHTS

- Day 1 had no significant activity, which could suggest either downtime, a lack of engagement, or a deliberate pause.

- Day 2 saw a dramatic surge in tweets during the afternoon, that triggered widespread engagement during that period.

- Day 3 maintained consistent tweeting, especially during early morning hours, indicating a sustained level of interest or activity before gradually tapering off.

---


--- Sentiment Analysis ---

```{r}
cleaned_tweets <- read.csv("cleaned_tweets.csv")

cleaned_tweets <- cleaned_tweets %>%
  mutate(sentiment_category = case_when(
    sentiment > 0 ~ "Positive",
    sentiment < 0 ~ "Negative",
    TRUE ~ "Neutral"
  ))

sentiment_by_device <- cleaned_tweets %>%
  group_by(tweetSource, sentiment_category) %>%
  summarise(count = n(), .groups = "drop")


sentiment_plot <- ggplot(sentiment_by_device, aes(x = sentiment_category, y = count, fill = sentiment_category)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ tweetSource) +
  labs(
    title = "Sentiment Analysis by Device",
    x = "Sentiment Category",
    y = "Count of Tweets"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("Positive" = "green", "Neutral" = "blue", "Negative" = "red"))


print(sentiment_plot)



```